### 2018-07-16 tables/linedepths.tex

%!TEX root = ../nir_companions.tex

\begin{table}
        \small
    \begin{threeparttable}[b]

   \caption{Contribution of synthetic lines within 2110--2160~nm of synthetic PHOENIX-ACES spectra to a binary model. \(F_{2}/F_{1}\) is the continuum flux ratio between a spectrum with the given \(T_{\textrm{eff}}\) and logg and a Sun-like spectrum with \(T_{\textrm{eff}}=5800\), logg = 4.5 (right most column). \emph{No.\ lines} is the number of spectral lines deeper than 5\% from the continuum of the individual spectra while \emph{avg.\ depth} is the mean depth of those lines. \emph{Cont.\ depth} is the average contribution, or depth, of these lines in the combined spectrum of a binary with a Sun-like spectrum. The SNR is signal-to-noise level required to have Gaussian noise \(\sigma\) =1/SNR equal to the \emph{cont.\ depth} level in the binary model. All synthetic spectra used here have [Fe/H]=0.0.}

    \begin{tabular}{*7c}
        \toprule
        \(T_{\textrm{eff}}\) (K)  & \multicolumn{2}{c}{2300} & \multicolumn{2}{c}{3200} & 5800 (\(\rm F_1\))\\
        logg & 5.0 & 4.5  & 5.0 & 4.5 & 4.5 \\
        \midrule
        \(F_2/F_1\) & 0.006 & 0.019 & 0.029  & 0.091 & 1.000 \\  
        % {>2\%}  & no. lines & 470 &  463 & 414  & 444 & 111 \\
        % & avg.\ depth & 0.20 & 0.23 & 0.10  & 0.12 & 0.04 \\
        % & cont. depth \tablefootmark{a} & 0.0012 &  0.0043 & 0.0028 &  0.0100 &  0.0333\tablefootmark{b} \\ 
        % \midrule
        no.\ lines & 464 & 463 & 365  & 413 & 31 \\
        avg.\ depth & 0.2  & 0.23& 0.11 & 0.12 & 0.10 \\
        cont.\ depth\tnote{a} &  0.0012 & 0.0043 &  0.0031 & 0.0100&  0.0833\tnote{b} \\ 
        SNR  & 833 & 232 & 323  & 100 & 12 \\ 
        %  SNR\(\rm _N\)  & 39 & 11 & 17  & 5 & 2 \\ 
        \bottomrule
    \end{tabular}
    \begin{tablenotes}
        \item [a] avg.\ depth \(\times~ \rm F_2 / (F_1 + F_2)\), where \(F_1\) is the component in the far right column.
        \item[b] avg.\ depth \(\times~ \rm F_1 / (F_1 + F_2)\), where \(\rm F_2\) is for the companion with \(T_{\textrm{eff}}\)=3200, logg=4.5.
    \end{tablenotes}
    \end{threeparttable}
    \label{tab:line_contributions}
\end{table}