### 2018-07-16 tables/observation_table.tex

%!TEX root = ../nir_companions.tex

% Table of observations
\begin{table*}
         \begin{threeparttable}[b]
           \centering    
            \caption{Details about the CRIRES observations times and settings. The estimated RV values for both component in each observation are provided. These are calculated using the best known orbital parameters and the companion mass \((M_{2}\sin{i}/M_{2})\). For hosts with multiple companions the RV value is for the largest companion only i.e. {HD 202206}B and {HD 168443}c. The RV difference between the host and the companion (\(RV_2 - RV_1\)) corresponds to the \(rv_2\) parameter in the binary model from Sect.~\ref{subsubsec:binary-model}.}
            \begin{tabular}{l c c c c r@{.}l r@{.}l r@{.}l}
                \toprule
                Object & Obs. \# & Start date  & Filter & Airmass  & \multicolumn{2}{c}{\(RV_1\)} & \multicolumn{2}{c}{\(RV_2\)} & \multicolumn{2}{c}{\(rv_2\)}  \\  % & \(Date \)
                &   & (yyyy-mm-dd hh:mm:ss)  &  & (at start) & \multicolumn{2}{c}{kms\(^{-1}\)} & \multicolumn{2}{c}{kms\(^{-1}\)} & \multicolumn{2}{c}{kms\(^{-1}\)}\\ % data ref    % & (JD\(^{\star} \))
                \midrule
                {HD 4747}   & 1 & 2012-07-06 07:36:06 & Ks     	      & 1.25  	  & $-$0    & 219 & $-$0  & 154 & 0&065 \\ %-1      & 2456114.81674
                {HD 162020} & 1 & 2012-07-04 06:23:22 & Ks     		& 1.30 		& $-$28  & 760 & 50 & 785\tnote{a}  & 79&545$^a$ \\ %-1      & 2456112.76624
                {HD 162020} & 2 & 2012-07-04 06:57:48 & Ks     		& 1.44  	& $-$28  & 717 & 48 & 440$^a$ & 77&157$^a$ \\ %-2      & 2456112.79015
                {HD 167665} & 1 & 2012-07-28 05:00:53 & Hx5e-2 	& 1.24 		& 7         & 581 & 18 & 024$^a$ & 10&443$^a$ \\ %-1a     & 2456136.70895
                {HD 167665} & 2 & 2012-07-28 05:37:27 & Hx5e-2 	& 1.39  	& 7         & 581 & 18 & 025$^a$  & 10&444$^a$ \\ %-1b     & 2456136.73434
                {HD 167665} & 3 & 2012-08-05 02:54:03 & Hx5e-2 	& 1.04  	& 7         & 575 & 18 & 163$^a$ & 10&588$^a$ \\ %-2      & 2456144.62087
                {HD 168443} & 1 & 2012-08-05 04:29:32 & Ks     		& 1.31 		& $-$0   & 121 & 50 & 932$^{a, b}$  & 51&053$^{a, b}$ \\ %-1      & 2456144.68718
                {HD 168443} & 2 & 2012-08-05 04:58:50 & Ks     		& 1.47 		& $-$0   & 121 & 51 & 189 $^{a, b}$ & 51&310$^{a, b}$ \\ %-2      & 2456144.70753
                {HD 202206} & 1 & 2012-07-12 06:54:44 & Ks     		& 1.01 		& 14      & 843 & 12 & 992$^b$  & -1&851 \\ %-1      & 2456120.78801
                {HD 202206} & 2 & 2012-07-13 05:41:40 & J       		  & 1.01 	  & 14      & 837 & 13 & 065$^b$  & -1&772 \\ %-2      & 2456121.73727
                {HD 202206} & 3 & 2012-07-11 08:29:55 & Ks     		& 1.15		& 14      & 849 & 12 & 920$^b$  & -1&929 \\ %-3      & 2456119.85411
                {HD 211847} & 1 & 2012-07-06 07:02:57 & Ks     		& 1.07 		& 6        & 613 & 7   & 171 & 0& 558\\ %-1      & 2456114.79372
                {HD 211847} & 2 & 2012-07-13 06:54:37 & Ks     		& 1.05 		& 6        & 614 & 7   & 167 & 0&553 \\ %-2      & 2456121.78793
                {HD 30501}  & 1 & 2012-04-07 00:08:29 & Hx5e-2 	 & 1.60 	 & 22      &  372 & 36 & 377 & 14&005 \\ %-1      & 2456024.50590
                {HD 30501}  & 2 & 2012-08-01 09:17:30 & Hx5e-2    & 1.42		 & 22      & 505 & 35  & 120 & 12&615 \\ %-2a     & 2456140.88716
                {HD 30501}  & 3 & 2012-08-02 08:47:30 & Hx5e-2 	 & 1.53 	 & 22      & 507 &  35 & 102 & 12&595 \\ %-3      & 2456141.86633
                {HD 30501}  & 4 & 2012-08-06 09:42:07 & Ks     		 & 1.28 	 & 22      & 514 & 35 & 031 & 12&517 \\ %-2b     & 2456145.90426
                \bottomrule

        \end{tabular}
        \begin{tablenotes}
          \item  [a] {$^a$}{Maximum RV given \(\textrm{M}_2\sin{i}\) only.}
          \item  {$^b$}{Largest mass companion only.}
        \end{tablenotes}
    \end{threeparttable}
    \label{tab:observations}
\end{table*}