### 2018-07-16 tables/uppermasslimits.tex

%!TEX root = ../nir_companions.tex
\begin{table}
    \begin{threeparttable}
        \centering
        \caption{Upper mass limits of target companions assuming a companion logg = 5.0. Masses are derived from \citet{baraffe_new_2015} evolutionary models using  \(T_{\textrm{eff}}\) and logg. The flux ratio \(\rm F_2/F_1\) is the absolute flux ratio between the cutoff temperature and the target host star.}
        \begin{tabular}{l c c c}
            \toprule
            Target & \(T_{\textrm{eff}}\) cutoff (K) & \(\rm F_2/F_1\) & Mass limit ($\rm M_{jup}$)\\
            \midrule
            {HD 4747}     &  3900 & 0.084 & 598\\ 
            {HD 162020} & 3900 & 0.147 & 598\\
            {HD 167665} & 3800 & 0.054 & 560\\
            {HD 168443} & 4000 & 0.094 & 618\\
            {HD 202206} & 3900 & 0.075 & 598\\
            {HD 211847} & 3900 & 0.079 & 598 \\
            {HD 30501}   & 3800\tnote{a} & 0.106 & 560\\
            \bottomrule
        \end{tabular}
        \begin{tablenotes}
                \item [a] {From observed spectra}
        \end{tablenotes}
    \end{threeparttable}
    \label{tab:mass_limits}
\end{table}