### 2018-07-26 tables/estimatedparameters.tex
Update Tables with more info.
Exp time, " exp, SNR.

%!TEX root = ../nir_companions.tex
\begin{table*}
  %   \begin{threeparttable}[b]
        \small
        \centering
        \caption{Estimated flux ratios, orbital semi-amplitude and RV separation of the companion, given the companion mass (\(\textrm{M}_{2}\) or \(\textrm{M}_{2} \sin{i}\)) from Table~\ref{tab:orbitparams} and observation times from Table~\ref{tab:observations}.} 
        \begin{tabular}{l | c c c c c c c | c c c}[hb]
            \toprule
            & Host& Companion &  Estimated  & Estimated &  Estimated & Estimated &  &    \\  % 2017
            Companion     & M$_{K}$& M$_{K}$& \(\rm F_{2}/F_{1} \)   & \(\rm N_{2}/N_{1} \) (noise ratio) & \(\rm K_2\) &   \(\Delta RV\) & Phase coverage \\
            & & & \textit{K}-band     & & (\kmps{}) & (\,ms\(^{-1}\)) & (\%) \\
            \midrule
            {HD 4747}        & & & \(3\times10^{-4} \)   & 76 &  -10.65 & -  &  -  \\  % 2017
            {HD 162020}   & & & \(7\times10^{-7} \)   & 1615  &  -98.92\tnote{a} &  2344.24     & 0.28~~  \\  %
            {HD 167665}    & & & \(2\times10^{-4} \)   &  105    &  -14.47\tnote{a}  &   138.45     & 0.18~~  \\  %  -- \(2\times10^{-5} \)  best case based on age rage.
            {HD 168443b} & & & \(1\times10^{-16} \)  &    \(1\times10^{8} \)   &  -64.65\tnote{a}&   257.16   & 0.035 \\ 
            {HD 168443c} & & &  \(1\times10^{-11} \)  &   \(4\times10^{5} \)     &  -18.05\tnote{a}  &   0.95   &  0.001 \\  %(c)
            {HD 202206}B  & & & \(8\times10^{-7} \)  &   1586 &  -6.79 & 145.17   & 0.74~  \\  %(B)   % May2017
            {HD 202206}c  & & &  \(5\times10^{-15}\)   &     \(2\times10^{7} \) &   -2.50     &   0.67     &  0.15~  \\  %(B)   % May2017
            {HD 211847}B  & & &  0.01 &  14   & $-$1.85 & 3.88   & 0.09~  \\  %B % 2017
            {HD 30501}      & & &  0.002  &  27  &  -16.12    &  1346.46      & 5.8~~  \\
            \bottomrule& & 
            \end{tabular}
    \begin{tablenotes}
        \item[a] {Maximum \(K_2\) only given \(M_2 \sin{i}\)}
      \end{tablenotes}
%    \end{threeparttable}
    \label{tab:estimatedparameters}
\end{table*}