### 2018-07-27 tables/observation_table.tex
Add number of artifacts to table

%!TEX root = ../nir_companions.tex

% Table of observations
\begin{table*}
       %  \begin{threeparttable}[b]
           \centering    
            \caption{Details about the each CRIRES observation. {\rd The time, settings, number of artefacts removed, the SNR obtained and the predicted orbital state of each system are provided.}}
            %\begin{tabular}{l c c c c cl cl r@{.}l r@{.}l r@{.}l}
            \begin{tabular}{l c c c c c c r@{.}l r@{.}l r@{.}l}
                \toprule
                Object & Obs. \# & Start date  & Filter & Airmass  & Artefacts & SNR & \multicolumn{2}{c}{\(RV_1\)} & \multicolumn{2}{c}{\(RV_2\)} & \multicolumn{2}{c}{\(rv_2\)}  \\  % & \(Date \)
                &   & (yyyy-mm-dd hh:mm:ss)  &  & (at start) & {\rd/ 32} & & \multicolumn{2}{c}{kms\(^{-1}\)} & \multicolumn{2}{c}{kms\(^{-1}\)} & \multicolumn{2}{c}{kms\(^{-1}\)}\\ % data ref    % & (JD\(^{\star} \))
                \midrule
                {HD 4747}   & 1 & 2012-07-06 07:36:06 & Ks     	      & 1.25  	  & 7 & 340 & $-$0    & 219 & $-$0  & 154 & 0&065 \\ %-1      & 2456114.81674
                {HD 162020} & 1 & 2012-07-04 06:23:22 & Ks     		& 1.30 		& 2 & 127 & $-$28  & 760 & 50 & 785\tnote{a}  & 79&545$^a$ \\ %-1      & 2456112.76624
                {HD 162020} & 2 & 2012-07-04 06:57:48 & Ks     		& 1.44  	& 2 & 128 & $-$28  & 717 & 48 & 440$^a$ & 77&157$^a$ \\ %-2      & 2456112.79015
                {HD 167665} & 1 & 2012-07-28 05:00:53 & Hx5e-2 	& 1.24 		& 7 & 371 & 7         & 581 & 18 & 024$^a$ & 10&443$^a$ \\ %-1a     & 2456136.70895
                {HD 167665} & 2 & 2012-07-28 05:37:27 & Hx5e-2 	& 1.39  	& 4 & 374 & 7         & 581 & 18 & 025$^a$  & 10&444$^a$ \\ %-1b     & 2456136.73434
                {HD 167665} & 3 & 2012-08-05 02:54:03 & Hx5e-2 	& 1.04  	& 4 & 358 & 7         & 575 & 18 & 163$^a$ & 10&588$^a$ \\ %-2      & 2456144.62087
                {HD 168443} & 1 & 2012-08-05 04:29:32 & Ks     		& 1.31 		& 2& 192 & $-$0   & 121 & 50 & 932$^{a, b}$  & 51&053$^{a, b}$ \\ %-1      & 2456144.68718
                {HD 168443} & 2 & 2012-08-05 04:58:50 & Ks     		& 1.47 		& 4 & 190 & $-$0   & 121 & 51 & 189 $^{a, b}$ & 51&310$^{a, b}$ \\ %-2      & 2456144.70753
                {HD 202206} & 1 & 2012-07-12 06:54:44 & Ks     		& 1.01 		& 3& 189 & 14      & 843 & 12 & 992$^b$  & -1&851 \\ %-1      & 2456120.78801
                {HD 202206} & 2 & 2012-07-13 05:41:40 & J       	  & 1.01 	  & 3 & 209 & 14      & 837 & 13 & 065$^b$  & -1&772 \\ %-2      & 2456121.73727
                {HD 202206} & 3 & 2012-07-11 08:29:55 & Ks     		& 1.15		& 4& 180 & 14      & 849 & 12 & 920$^b$  & -1&929 \\ %-3      & 2456119.85411
                {HD 211847} & 1 & 2012-07-06 07:02:57 & Ks     		& 1.07 		& 4& 272 & 6        & 613 & 7   & 171 & 0& 558\\ %-1      & 2456114.79372
                {HD 211847} & 2 & 2012-07-13 06:54:37 & Ks     		& 1.05 		& 5& 283 & 6        & 614 & 7   & 167 & 0&553 \\ %-2      & 2456121.78793
                {HD 30501}  & 1 & 2012-04-07 00:08:29 & Hx5e-2 	 & 1.60 	 & 3& 217 & 22      &  372 & 36 & 377 & 14&005 \\ %-1      & 2456024.50590
                {HD 30501}  & 2 & 2012-08-01 09:17:30 & Hx5e-2    & 1.42     & 10& 212 & 22      & 505 & 35  & 120 & 12&615 \\ %-2a     & 2456140.88716
                {HD 30501}  & 3 & 2012-08-02 08:47:30 & Hx5e-2 	 & 1.53 	 & 8& 237 & 22      & 507 &  35 & 102 & 12&595 \\ %-3      & 2456141.86633
                {HD 30501}  & 4 & 2012-08-06 09:42:07 & Ks     		 & 1.28 	 & 7& 235& 22      & 514 & 35 & 031 & 12&517 \\ %-2b     & 2456145.90426
                \bottomrule
                & & & & 
        \end{tabular}
        \begin{tablenotes}
          \item  [a] {$^a$}{Maximum RV given \(\textrm{M}_2\sin{i}\) only.}
          \item  {$^b$}{Largest mass companion only.}
        \end{tablenotes}
 %   \end{threeparttable}
    \label{tab:observations}
\end{table*}