### 2018-08-02 tables/stellarparameters.tex

%!TEX root = ../nir_companions.tex

%% Table of stellar parameters
\begin{table*}
        \small
        \centering
        \begin{threeparttable}[b]

        \caption{Stellar parameters of the target companion's host stars. V is the apparent magnitude taken from {SIMBAD}~\citep{wenger_simbad_2000}. {\rd Distances were calculated from the GAIA parallax measurements.}}
        \begin{tabular}{l c c r@{$~\pm~$}l r@{$~\pm~$}l r@{$~\pm~$}l r@{$~\pm~$}l c c c}
            \toprule
            Star & SpT & V &  \multicolumn{2}{c}{\(T_{\textrm{eff}}\) (K)} &  \multicolumn{2}{c}{logg (cm s\(^{-2} \))}  & \multicolumn{2}{c}{[Fe/H]} &  \multicolumn{2}{c}{\(M_1\) (M\(_{\sun} \))} & Age (Gyr) & d (pc) & Reference\\
            \midrule
            {HD 4747}     & K0V & 7.15 & 5316 & 50 & 4.48 & 0.10  & $-$0.21 & 0.05 & 0.81 & 0.02  & $3.3 \pm 2.3$   & $18.80 \pm 0.04$ & 1, 2, 3, 8 \\ 
            {HD 162020} & K3V & 9.12 & 4723 & 71 & 4.31 & 0.18  & $-$0.10 & 0.03 & 0.74 & 0.07  & $3.1 \pm 2.7$   & $30.85 \pm 0.06$ & 4, 5, 6, 8 \\  
            {HD 167665} & F9V & 6.40 & 6224 & 50 & 4.44 & 0.10  & 0.05       & 0.06 & 1.14 & 0.03  & 0.7 -- 3.6           & $ 31.24 \pm 0.06$ & 1, 8 \\
            {HD 168443} & G6V & 6.92 & 5617 & 35 & 4.22 & 0.05 & 0.06       & 0.05 & 1.01 & 0.07  & $10.0 \pm 0.3$  & $39.67 \pm 0.12$ & 5, 6, 8 \\ 
            {HD 202206} & G6V & 8.07 & 5757 & 25 & 4.47 & 0.03 & 0.29       & 0.02 & 1.04 & 0.07  & $2.9 \pm 1.0$    & $46.03 \pm 0.14$ & 5, 7, 8 \\ 
            {HD 211847} & G5V & 8.62 & 5715 & 24 & 4.49 & 0.05 & $-$0.08 & 0.02 & 0.92 & 0.07   & 0.1 -- 6.0            & $48.81 \pm 0.13 $ & 1, 2, 4, 8 \\ 
            {HD 30501}   & K2V & 7.59  & 5223 & 50 & 4.56 & 0.10 & 0.06       & 0.06 & 0.81 & 0.02  & 0.8 -- 7.0           & $20.37 \pm 0.01$ & 1, 4, 8 \\
            \bottomrule
        \end{tabular}
         \label{tab:starparams}
        \begin{tablenotes}
           \item[] References: (1)~\citet{sahlmann_search_2011}; (2)~\citet{santos_spectroscopic_2005}; (3)~\citet{crepp_trends_2016}; (4)~\citet{tsantaki_deriving_2013}; (5)~\cite{bonfanti_age_2016}; (6)~\citet{santos_spectroscopic_2004}; (7)~\citet{sousa_spectroscopic_2008}; {\rd(8)}~\citet{collaboration_gaia_2018};
         \end{tablenotes}
       \end{threeparttable}

\end{table*}