### 2018-08-17 tables/estimatedparameters.tex
type setting Msun and (mag) label

%!TEX root = ../nir_companions.tex
\begin{table*}
  % 
        \small
        \centering
       \begin{threeparttable}[b]
        \caption{Estimated flux ratios, orbital semi-amplitude and RV separation of the companion, given the companion mass (\(\textrm{M}_{2}\) or \(\textrm{M}_{2} \sin{i}\)) from Table~\ref{tab:orbitparams} and observation times from Table~\ref{tab:observations}.} 
        \begin{tabular}{l c c c c c c c c c c}%[hb]
            \toprule
            & Host& Companion & Estimated & Estimated & Estimated & Estimated & & \\  % 2017
            Companion & M$_{K}$ & M$_{K}$ & \(\rm F_{2}/F_{1} \) & \(\rm N_{2}/N_{1} \) (noise ratio) & \(\rm K_2\) & \(\Delta RV\) & Phase coverage\\
            & (mag) & (mag)& \textit{K}-band & & (\kmps{}) & (\,ms\(^{-1}\)) & (\%)\\
            \midrule
            {HD 4747} & 3.82 & 14.17 & \(7\times10^{-5} \) & 76 & -10.65 & -- & --\\  % 2017
            {HD 162020} & 4.10 & 23.36 & \(2\times10^{-8} \) & 1615 & -98.92\tnote{a} & 2344.24 & 0.28\\  %
            {HD 167665} & 2.60 & 13.21 & \(6\times10^{-5} \) & 105 & -14.47\tnote{a} & 138.45 & 0.18\\  %  -- \(2\times10^{-5} \)  best case based on age rage.
            {HD 168443b} & 2.35 & 42.19 & \(1\times10^{-16} \) & \(1\times10^{8} \) & -64.65\tnote{a}& 257.16 & 0.035\\ 
            {HD 168443c} & 2.35 & 29.55 & \(1\times10^{-11} \) & \(4\times10^{5} \) & -18.05\tnote{a} & 0.95 & 0.001\\  %(c)
            {HD 202206}B & 3.04& 21.63 & \(4\times10^{-8} \) & 1586 & -6.79 & 145.17 & 0.74\\  %(B)   % May2017
            {HD 202206}c & 3.04& 45.63 & \(9\times10^{-18}\) & \(2\times10^{7} \) & -2.50 & 0.67 & 0.15\\  %(B)   % May2017
            {HD 211847}B & 3.50 & 8.40 & 0.011 & 14 & $-$1.85 & 3.88 & 0.09\\  %B % 2017
            {HD 30501} & 3.96 & 10.38 & 0.003 & 27 & -16.12 & 1346.46 & 5.8\\
            \bottomrule& & 
            \end{tabular}
            \label{tab:estimatedparameters}
    \begin{tablenotes}
        \item[a] {Maximum \(K_2\) only given \(M_2 \sin{i}\)}
      \end{tablenotes}
  \end{threeparttable}

\end{table*}