### 2018-08-17 tables/linedepths.tex

%!TEX root = ../nir_companions.tex
\begin{table}
    \small
    \centering
    \begin{threeparttable}[b]

   \caption{Contribution of synthetic lines within 2110--2160~nm of synthetic PHOENIX-ACES spectra to a binary model. \(F_{2}/F_{1}\) is the continuum flux ratio between a spectrum with the given \(T_{\textrm{eff}}\) and logg and a Sun-like spectrum with \(T_{\textrm{eff}}=5800\), logg = 4.5 (right most column). \emph{No.\ lines} is the number of spectral lines deeper than 5 per cent from the continuum of the individual spectra while \emph{avg.\ depth} is the mean depth of those lines. \emph{Cont.\ depth} is the average contribution, or depth, of these lines in the combined spectrum of a binary with a Sun-like spectrum. The SNR is signal-to-noise level required to have Gaussian noise \(\sigma\) = 1/SNR equal to the \emph{cont.\ depth} level in the binary model. All synthetic spectra used here have [Fe/H] = 0.0.}

    \begin{tabular}{*7c}
        \toprule
        \(T_{\textrm{eff}}\) (K)  & \multicolumn{2}{c}{2300} & \multicolumn{2}{c}{3200} & 5800 (\(\rm F_1\))\\
        logg & 5.0 & 4.5  & 5.0 & 4.5 & 4.5 \\
        \midrule
        \(F_2/F_1\) & 0.006 & 0.019 & 0.029  & 0.091 & 1.000 \\
        no.\ lines & 464 & 463 & 365  & 413 & 31 \\
        avg.\ depth & 0.2  & 0.23& 0.11 & 0.12 & 0.10 \\
        cont.\ depth\tnote{a} &  0.0012 & 0.0043 &  0.0031 & 0.0100&  0.0833\tnote{b} \\
        SNR  & 833 & 232 & 323  & 100 & 12 \\
        \bottomrule
    \end{tabular}\label{tab:line_contributions}
    \begin{tablenotes}
        \item [a] avg.\ depth \(\times~ \rm F_2 / (F_1 + F_2)\), where \(F_1\) is the component in the far right column.
        \item[b] avg.\ depth \(\times~ \rm F_1 / (F_1 + F_2)\), where \(\rm F_2\) is for the companion with \(T_{\textrm{eff}}\) = 3200, logg = 4.5.
    \end{tablenotes}
  \end{threeparttable}
\end{table}