### 2018-08-17 tables/phoenixparameters.tex

%!TEX root = ../nir_companions.tex

\begin{table}
    \centering
    \caption{Full parameter space of the PHOENIX-ACES spectral grid.}
    \begin{tabular}{lr@{ -- }lc}  % Seperate columns with --
        \toprule
        & \multicolumn{2}{c}{Range}       & Step size\\
        \midrule
        \ \(T_{\textrm{eff}}\) [K] &  2300 & 7000  & 100 \\
        &  7000 & 12000 & 200 \\
        \  logg     &  0.0 & +6.0   & 0.5 \\
        \ [Fe/H]   &  -4.0 & $-$2.0  & 1.0 \\
        &  -2.0 & +1.0  & 0.5 \\
        \  \(\alpha\)/Fe &  -0.2 & +1.2  & 0.2 \\
        \bottomrule
    \end{tabular} \label{tab:phoenix}
\end{table}